### 2014-06-12 src/paper1.tex
Updated paper 1

@@

\noindent Published in the Journal of Neuroscience \cite{clemens2011}.

@@

\section*{Abstract}
Most evidence that the brain uses Bayesian inference to integrate noisy sensory signals optimally has been obtained by showing that the noise levels in each modality separately can predict performance in combined conditions. Such a forward approach is difficult to implement when the various signals cannot be measured in isolation, as in spatial orientation, which involves the processing of visual, somatosensory, and vestibular cues. Instead, we applied an inverse probabilistic approach, based on optimal observer theory. Our goal was to investigate whether the perceptual differences found when probing two different states ---body-in-space and head-in-space orientation--- can be reconciled by a shared scheme using all available sensory signals. Using a psychometric approach, seven human subjects were tested on two orientation estimates at tilts \textless 120\textdegree: perception of body tilt [subjective body tilt (SBT)] and perception of visual vertical [subjective visual vertical (SVV)]. In all subjects, the SBT was more accurate than the SVV, which showed substantial systematic errors for tilt angles beyond 60\textdegree. Variability increased with tilt angle in both tasks, but was consistently lower in the SVV. The sensory integration model fitted both datasets very nicely. A further experiment, in which supine subjects judged their head orientation relative to the body, independently confirmed the predicted head-on-body noise by the model. Model predictions based on the derived noise properties from the various modalities were also consistent with previously published deficits in vestibular and somatosensory patients. We conclude that Bayesian computations can account for the typical differences in spatial orientation judgments associated with different task requirements.

@@

\begin{figure}
    %\includegraphics[width=0.75\textwidth]{src/paper1/figure1.eps}
\end{figure}

@@

\begin{figure}
    %\includegraphics[width=0.5\textwidth]{src/paper1/figure2.eps}
\end{figure}

@@

\begin{equation}
\label{p1psych}
P(x) = \lambda + (1 - 2\lambda) \frac{1}{\sigma \sqrt{2\pi}} \int_{-\infty}^{x}{e^{-(y-\mu)^2 / 2\sigma^2}}dy,
\end{equation}
% Equation 1

@@

\begin{equation}
\tilde{B}_S = w_{BD} \cdot \hat{B}_S + w_{BI} \cdot (\hat{H}_S - \hat{H}_B),
\end{equation}

@@

\begin{equation}
w_{BD} = \frac{1/\sigma^2_{BS}}{1/(\sigma^2_{HS} + \sigma^2_{HB}) + 1/\sigma^2_{BS}}
\end{equation}

@@

\begin{equation}
w_{BI} = \frac{1/(\sigma^2_{HS} + \sigma^2_{HB})}{1/(\sigma^2_{HS} + \sigma^2_{HB}) + 1/\sigma^2_{BS}}
\end{equation}

@@

\begin{equation}
\sigma^2(\tilde{B}_S) = \frac{
\sigma^2_{BS} \cdot (\sigma^2_{HS} + \sigma^2_{HB})
}{
\sigma^2_{BS} + (\sigma^2_{HS} + \sigma^2_{HB})
}
\end{equation}

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper1/figure3.eps}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=1.0\textwidth]{src/paper1/figure4.eps}
\end{figure}

@@

% Table 1

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper1/figure5.eps}
\end{figure}

@@

% Table 2

@@

\begin{figure}
    \includegraphics[width=0.6\textwidth]{src/paper1/figure6.eps}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper1/figure7.eps}
\end{figure}

@@

\begin{figure}
%    \includegraphics[width=1.0\textwidth]{src/paper1/figure8.eps}
\end{figure}

That the variance of the peak locations across multiple trials, s2(H~S), is smaller can be shown by applying the rules of noise propagation to Equation 13.

>>>>

\small {\bf Abstract} Most evidence that the brain uses Bayesian inference to integrate noisy sensory signals optimally has been obtained by showing that the noise levels in each modality separately can predict performance in combined conditions. Such a forward approach is difficult to implement when the various signals cannot be measured in isolation, as in spatial orientation, which involves the processing of visual, somatosensory, and vestibular cues. Instead, we applied an inverse probabilistic approach, based on optimal observer theory. Our goal was to investigate whether the perceptual differences found when probing two different states ---body-in-space and head-in-space orientation--- can be reconciled by a shared scheme using all available sensory signals. Using a psychometric approach, seven human subjects were tested on two orientation estimates at tilts \textless 120\textdegree: perception of body tilt [subjective body tilt (SBT)] and perception of visual vertical [subjective visual vertical (SVV)]. In all subjects, the SBT was more accurate than the SVV, which showed substantial systematic errors for tilt angles beyond 60\textdegree. Variability increased with tilt angle in both tasks, but was consistently lower in the SVV. The sensory integration model fitted both datasets very nicely. A further experiment, in which supine subjects judged their head orientation relative to the body, independently confirmed the predicted head-on-body noise by the model. Model predictions based on the derived noise properties from the various modalities were also consistent with previously published deficits in vestibular and somatosensory patients. We conclude that Bayesian computations can account for the typical differences in spatial orientation judgments associated with different task requirements.

@@

\noindent\underline{ \hspace{4cm} }

\noindent This chapter has been published as \newline
\noindent {\bf Clemens, I.A.H.}, De Vrijer, M., Selen, L.P.J., Van Gisbergen, J.A.M. and Medendorp, W.P. (2011). Multisensory processing in spatial orientation: An inverse probabilistic approach. \emph{Journal of Neuroscience}, 31(14): 5365-5377. \newline
%\noindent Authors Clemens and De Vrijer contributed equally to this work.

@@

@@

\begin{figure}
    \includegraphics[width=0.75\textwidth]{src/paper1/figure1.pdf}

    \caption{Schematic representation of the sensory integration model. Sensory signals, denoted by a hat symbol ($\hat{ }$), are assumed to be calibrated accurately, but contaminated by Gaussian noise. Optimal estimates are denoted by a tilde ($\tilde{ }$). Body sensors, neck sensors, and otoliths provide information about orientation of body in space ($B_S$), head on body ($H_B$), and head in space ($H_S$), respectively. Neck signal ($\hat{H}_B$) is used for a reference frame transformation of otolith information into a body-in-space signal ($\hat{H}_S - \hat{H}_B = \hat{B}_{SI}$), and for a transformation of body-tilt information into a head-in-space signal ($\hat{B}_S + \hat{H}_B = \hat{H}_{SI}$). For an optimal estimate of body-in-space orientation, $\tilde{B}_S$ (SBT task), the model combines the body-sensor signal ($\hat{B}_S$, red pathway) with a reference-frame-transformed otolith signal ($\hat{B}_{SI}$, green pathway). Relative contributions of the two pathways ($w_{BD}$ and $w_{BI}$) depend on their relative precision (Eq. \ref{p1eqn2}). The scheme shows a symmetrical arrangement with two priors, but there is ample reason to believe that their effects are not identical. The simplest explanation of current and previous SBT data (see Materials and Methods, SBT computation) indicates that the associated prior in this task is uniform, which implies that $w_{BP}$ can be ignored. In the SVV task, an optimal estimate of head-in-space ($\tilde{H}_S$) is obtained by integration of otolith information ($\hat{H}_S$, green pathway), reference-frame-transformed information from body sensors ($\hat{H}_{SI}$, red pathway), and a significant contribution from prior information ($H_{SP}$, blue pathway). Relative weights are denoted by $w_{HD}$, $w_{HI}$, and $w_{HP}$, respectively. Estimate of line-in-space orientation is obtained by combining $\tilde{H}_S$ and estimates of eye-in-head ($\tilde{E}_H$) and line-on-eye ($\tilde{L}_E$) orientation. Noise variance in body sensors ($\sigma^2_{BS}$), neck sensors ($\sigma^2_{HB}$), otoliths ($\sigma^2_{HS}$), and width of prior ($\sigma^2_{HSP}$) defines their relative weights (see Materials and Methods). Otolith noise may depend on tilt angle (Eq. \ref{p1eqn11}). Note that the process of sensory integration, denoted here by summation of weighted sensory signals, is equivalent to multiplication of the underlying probability distributions (Eqs. \ref{p1eqn2} and \ref{p1eqn6} and Appendix).}
\end{figure}

@@

%\begin{figure}
\begin{wrapfigure}{l}[5pt]{0.5\textwidth}
    \includegraphics[width=0.5\textwidth]{src/paper1/figure2.pdf}

    \caption{Tilt paradigm in $SBT_{90}$ task. T1, T2, T3, Test angles at which the subject was prompted with a beep signal (*) to indicate whether body orientation was CW or CCW from the instructed reference orientation (i.e., 90\textdegree in this example). D1, D2, Detour angles randomly drawn from detour range (30 Ð 40\textdegree CW and CCW from center of test range). Rotations from detour (D) to test (T) angle were performed in a noisy fashion (see Materials and Methods, SBT).}
%\end{figure}
\end{wrapfigure}

@@

\begin{equation}
\label{p1eqn1}
P(x) = \lambda + (1 - 2\lambda) \frac{1}{\sigma \sqrt{2\pi}} \int_{-\infty}^{x}{e^{-(y-\mu)^2 / 2\sigma^2}}dy,
\end{equation}
% Equation 1

@@

\begin{equation}
\label{p1eqn2}
\tilde{B}_S = w_{BD} \cdot \hat{B}_S + w_{BI} \cdot (\hat{H}_S - \hat{H}_B),
\end{equation}

@@

\begin{equation}
\label{p1eqn3}
w_{BD} = \frac{1/\sigma^2_{BS}}{1/(\sigma^2_{HS} + \sigma^2_{HB}) + 1/\sigma^2_{BS}}
\end{equation}

@@

\begin{equation}
\label{p1eqn4}
w_{BI} = \frac{1/(\sigma^2_{HS} + \sigma^2_{HB})}{1/(\sigma^2_{HS} + \sigma^2_{HB}) + 1/\sigma^2_{BS}}
\end{equation}

@@

\begin{equation}
\label{p1eqn5}
\sigma^2(\tilde{B}_S) = \frac{
\sigma^2_{BS} \cdot (\sigma^2_{HS} + \sigma^2_{HB})
}{
\sigma^2_{BS} + (\sigma^2_{HS} + \sigma^2_{HB})
}
\end{equation}

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper1/figure3.pdf}

    \caption{SBT versus SVV performance in one subject (S1). Top, SBT. Proportion of CW responses is plotted against body orientation relative to the reference orientation (0\textdegree, \textpm45\textdegree, or \textpm90\textdegree). $\mu > 0\degree$ indicates tilt underestimation. Bottom, SVV. Proportion of CW responses is plotted against line orientation relative to vertical. Solid lines, Best-fit cumulative Gaussians, typified by $\mu$ and $\sigma$.}  
\end{figure}

@@

\begin{figure}
	\includegraphics[width=1.0\textwidth]{src/paper1/figure4.pdf}

    \caption{Model predictions superimposed on parameters from the psychometric fits to the SBT (two top rows) and SVV data (two bottom rows). Accuracy and variability characteristics as a function of roll-tilt angle are shown; values are psychometric fits ($\mu$ and $\sigma$ values, [CIRCLES]) and model predictions (line) from all subjects. Mean data and mean predictions across subjects are plotted in the rightmost column.}
\end{figure}

@@

\begin{table}

\begin{tabular}{lllllll}
\hline
Subject & $a_{HS}$ (\textdegree/\textdegree) & $b_{HS}$ (\textdegree) & $\sigma_{HSP}$ (\textdegree) & $\sigma_{BS}$ (\textdegree) & $\sigma_{HB}$ (\textdegree) & $A_{OCR}$ (\textdegree) \\
\hline
S1 & 0.23 \textpm 0.02 & 1.2 \textpm 0.32 & 11.6 \textpm 1.0 & 12.3 \textpm 1.1 & 3.3 \textpm 1.2 & 27.0 \textpm 2.2 \\
S2 & 0.12 \textpm 0.02 & 1.2 \textpm 0.52 & 9.4 \textpm 1.1 & 8.4 \textpm 2.9 & 6.4 \textpm 4.1 & 17.0 \textpm 3.8 \\
S3 & 0.20 \textpm 0.03 & 1.1 \textpm 0.42 & 14.4 \textpm 1.7 & 6.7 \textpm 1.9 & 9.3 \textpm 2.4 & 17.5 \textpm 2.1 \\
S4 & 0.07 \textpm 0.50 & 3.9 \textpm n/a & 11.2 \textpm 1.3 & 12.6 \textpm 2.3 & 7.1 \textpm 3.5 & 0 \textpm n/a \\
S5 & 0.11 \textpm n/a & 3.3 \textpm 1.0 & 18.7 \textpm 4.8 & 15.0 \textpm n/a & 3.6 \textpm 2.1 & 1.06 \textpm n/a \\
S6 & 0.23 \textpm 0.09 & 3.0 \textpm 1.5 & 9.5 \textpm 1.1 & 8.0 \textpm 0.83 & 1.8 \textpm n/a & 18.8 \textpm 4.1 \\
S7 & 0.20 \textpm 0.14 & 3.2 \textpm 1.0 & 12.8 \textpm 2.4 & 12.7 \textpm 6.1 & 3.0 \textpm n/a & 20.8 \textpm 9.0 \\
Mean & 0.16 \textpm 0.06 & 2.4 \textpm 1.2 & 12.5 \textpm 3.2 & 10.8 \textpm 3.1 & 4.9 \textpm 2.7 & 14.6 \textpm 10.2 \\
\hline
\multicolumn{7}{l}{Imposed fit limits were as follows: $a_{HS}$: 0.5\textdegree/\textdegree; $b_{HS}$, $\sigma_{HSP}$, $\sigma_{BS}$, $\sigma_{HB}$, 50\textdegree; $A_{OCR}$, 30\textdegree. SD values are not shown (n/a) when bootstrapped values formed a skewed distribution. $a_{HS}$, Tilt-related increase in otolith noise; $b_{HS}$, otolith noise in upright position; $\sigma{HSP}$, width of head-tilt prior; $\sigma{BS}$, noise in body-in-space sensors; $\sigma{HB}$, noise in neck sensors; $A_{OCR}$, uncompensated ocular counterroll.} \\
\end{tabular}

\caption{Best-fit parameter and bootstrap-based SD values}
\end{table}

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper1/figure5.pdf}

    \caption{Tilt dependence of weight factors in SBT (top) and SVV (bottom) for each subject. Trends with tilt angle are similar for all subjects. Head-in-space prior is only involved in SVV computations. Means across subjects are plotted in the rightmost column.}
\end{figure}

@@

\begin{table}
\begin{tabular}{llllll}
\hline
Subject & MLE full model & MLE reduced model & $p$ & BIC full model & BIC psychometric fits \\
\hline
S1 & 231.5 & 312.4 & SMALLERTHAN 0.001 & 498.1 & 540.3 \\
S2 & 197.8 & 216.9 & SMALLERTHAN 0.001 & 430.6 & 539.3 \\
S3 & 267.5 & 313.4 & SMALLERTHAN 0.001 & 570.0 & 548.2 \\
S4 & 207.5 & 218.2 & SMALLERTHAN 0.001 & 450.0 & 569.1 \\
S5 & 195.6 & 216.3 & SMALLERTHAN 0.001 & 426.2 & 571.0 \\
S6 & 163.5 & 183.2 & SMALLERTHAN 0.001 & 362.0 & 502.5 \\
S7 & 248.0 & 260.0 & SMALLERTHAN 0.001 & 531.1 & 633.5 \\ 
\hline
\multicolumn{6}{l}{Log-likelihood ratio test of the full model (with indirect pathways) and reduced model (without indirect pathways). For all subjects, the full model outperforms the reduced model lacking multisensory integration through the indirect pathways. BIC values are much lower for the Bayesian integration model compared with separate psychometric fits in six of seven subjects.} \\
\end{tabular}
\caption{Validation of the model}
\end{table}

@@

%\begin{figure}
\begin{wrapfigure}{l}[5pt]{0.5\textwidth}
    \includegraphics[width=0.5\textwidth]{src/paper1/figure6.pdf}
%\end{figure}
\end{wrapfigure}

@@

%\begin{figure}
\begin{wrapfigure}{l}[5pt]{0.5\textwidth}
    \includegraphics[width=0.5\textwidth]{src/paper1/figure7.pdf}
\end{wrapfigure}
%\end{figure}

@@

%\begin{figure}
\begin{wrapfigure}{l}[5pt]{0.75\textwidth}
    \includegraphics[width=0.75\textwidth]{src/paper1/figure8.pdf}
%\end{figure}
\end{wrapfigure}

That the variance of the peak locations across multiple trials, $\sigma^2(\tilde{H}_S)$, is smaller can be shown by applying the rules of noise propagation to Equation 13.

### 2014-06-12 src/paper2.tex
Added some figures.
Updated style
Updated paper 1

@@

\section*{Abstract}
In order to maintain visual stability during self-motion, the brain needs to update any ego-centric spatial representations of the environment. Here, we use a novel psychophysical approach to investigate how, and to which extent, the brain integrates visual, extraocular, and vestibular signals pertaining to this spatial update. Participants were oscillated sideways at a frequency of 0.63 Hz while keeping gaze fixed on a stationary light. When the motion direction changed, a reference target was shown either in front or behind the fixation point. At the next reversal, half a cycle later, we tested updating of this reference location by asking participants to judge whether a briefly flashed probe was shown to the left or right of the memorized target. We show that updating is not only biased, but that the direction and magnitude of this bias depend on both gaze and object location, implying that a gaze-centered reference frame is involved. Using geometric modeling, we further show that the gaze-dependent errors can be caused by an underestimation of translation amplitude, by a bias of visually perceived objects towards the fovea (i.e., a foveal bias), or by a combination of both.

@@

% Figure 1

@@

% Figure 2

@@

% Figure 3

@@

% Figure 4

@@

% Figure 5

>>>>

\thispagestyle{empty}

@@

\vfill

\section*{Abstract}
\small
In order to maintain visual stability during self-motion, the brain needs to update any ego-centric spatial representations of the environment. Here, we use a novel psychophysical approach to investigate how, and to which extent, the brain integrates visual, extraocular, and vestibular signals pertaining to this spatial update. Participants were oscillated sideways at a frequency of 0.63 Hz while keeping gaze fixed on a stationary light. When the motion direction changed, a reference target was shown either in front or behind the fixation point. At the next reversal, half a cycle later, we tested updating of this reference location by asking participants to judge whether a briefly flashed probe was shown to the left or right of the memorized target. We show that updating is not only biased, but that the direction and magnitude of this bias depend on both gaze and object location, implying that a gaze-centered reference frame is involved. Using geometric modeling, we further show that the gaze-dependent errors can be caused by an underestimation of translation amplitude, by a bias of visually perceived objects towards the fovea (i.e., a foveal bias), or by a combination of both.

\noindent Published in the Journal of Vision \cite{clemens2012}.

\newpage

@@

\begin{figure}
	\includegraphics[width=1.0\textwidth]{src/paper2/figure1.pdf}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper2/figure2.pdf}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.75\textwidth]{src/paper2/figure3.pdf}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper2/figure4.pdf}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper2/figure5.pdf}
\end{figure}
